---
title: "Rarecoal model fit"
output:
    html_document:
        code_folding: hide
params:
  filename_template: "~/Data/NaDene/revision/models/EUR_SEA_SIB_CHK_SAM.model1.template.txt"
  filename_params: "~/Data/NaDene/revision/rarecoal-results/EUR_SEA_SIB_CHK_SAM/model1_core1.m4.maxl.paramEstimates.txt"
  filename_fitTable: "~/Data/NaDene/revision/rarecoal-results/EUR_SEA_SIB_CHK_SAM/model1_core1.m4.maxl.summaryFitTable.txt"
---

# Loading Libraries and Data
Here we load some necessary libraries and the specified files.

Template-file: `r params$filename_template`

Params-file: `r params$filename_params`

fitTable-file: `r params$filename_fitTable`

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(knitr)
filename_template <- params$filename_template
filename_params <- params$filename_params
filename_fitTable <- params$filename_fitTable
```

```{r loading template and results, message = FALSE, warning = FALSE}
template_str <- read_file(filename_template)
# r = "BRANCHES\\s*=\\s*\\[\\s*[:alnum:]+[,\\s*[:alnum:]]\\]\\s*.*"
r <- "^BRANCHES\\s*=\\s*\\[(.+)\\]\\s*EVENTS\\s*=\\s*\\[([\\S\\s]*)\\]"
m <- str_match(template_str, r)
branches_str <- m[1,2]
events_str <- m[1,3]
branches <- str_split(branches_str, ",[\\s*]")[[1]]
events <- str_squish(str_split(events_str, ";[\\s*]")[[1]])

f <- file(filename_params, "r")
first_line <- readLines(f, n = 1)
close(f)

if(startsWith(first_line, "#")) {
  paramsDat <- read_tsv(filename_params, comment = "#")
} else {
  paramsDat <- read_tsv(filename_params, col_names = c("Parameter", "Estimate"))
}

for (i in 1:nrow(paramsDat))
  events <- str_replace(events, str_c("<", paramsDat[[i,1]], ">"), toString(paramsDat[[i,2]]))
summaryFitData <- read_tsv(filename_fitTable) %>%
  mutate(Populations=factor(Populations, level = unique(Populations)))
```

The following table shows the maximum likelihood parameter estimates for the model

```{r viewing parameter estimates}
paramsDat %>% mutate(Estimate=ifelse(startsWith(Parameter, "p"), format(round(Estimate*20000), big.mark = ",", scientific = FALSE),
    ifelse(startsWith(Parameter, "t"), format(round(Estimate*29*40000), big.mark = ",", scientific = FALSE),
      ifelse(startsWith(Parameter, "adm"), str_c(round(Estimate*100, 1), "%"), format(round(Estimate), big.mark = ","))))) %>% kable()
```

Here is a tree plot of the model. Times are indicated by y-coordinate, and admixture proportions are written as labels.

```{r tree plot}
plot.new()
n <- length(branches)
event_fields <- events %>% str_split("\\s")
max_time <- (event_fields %>% map(function(x) as.double(x[2])) %>% reduce(max)) * 40000 * 29
plot.window(c(1,n), c(-(max_time * 0.05),max_time*1.05))
for (e in event_fields) {
  if(e[1] %in% c("J", "K")) {
    time <- as.double(e[2]) * 40000 * 29
    to_pop <- e[3]
    from_pop <- e[4]
    to_i <- which(to_pop == branches)
    from_i <- which(from_pop == branches)
    lines(c(to_i, to_i, from_i, from_i), c(0, time, time, 0), lwd=2)
  }
  if(e[1] == "S") {
    time <- as.double(e[2]) * 40000 * 29
    to_pop <- e[3]
    from_pop <- e[4]
    frac <- as.double(e[5])
    to_i <- which(to_pop == branches)
    from_i <- which(from_pop == branches)
    arrows(to_i, time, from_i, time, 0.1, col="blue", lty="dashed")
    x_adj <- if(to_i < from_i) 1 else -0.2
    text(from_i, time, labels=str_c(toString(round(frac * 100, digits=1)), "%"), adj=c(x_adj, -0.5), col="blue")
  }
}
text(1:n, rep(0, n), labels=branches, adj=c(0.5, 2))
axis(2)
```

Here is a fit of the model based on rare allele sharing statistics

```{r fit plot}
summaryFitData %>%
  transmute(Populations, real=AlleleSharing, predicted=Predicted) %>%
  gather(real, predicted, key="type", value="ras") %>%
  ggplot() +
    geom_bar(aes(x=Populations, y=ras*1e9, fill=type), stat="identity", position=position_dodge()) +
    scale_y_log10() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

And here is the relative deviation between model and fit:

```{r fit deviation plot}
ggplot(summaryFitData, aes(x=Populations, y=`relDev%`)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```